
\documentclass[smallextended]{svjour3}
 
\usepackage[authoryear,square]{natbib} 

%no bookmarks - conf' requirements.
\usepackage[bookmarks=false]{hyperref}
\pdfminorversion=5

\usepackage{array}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{pdfpages}

%\usepackage{cite}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\include{pslatex}

\usepackage{floatrow}
\usepackage{fancyvrb}
\usepackage[flushleft]{threeparttable}

\usepackage{framed}
\makeatletter
\renewenvironment{framed}{%
 \def\FrameCommand##1{\hskip\@totalleftmargin
 \fboxsep=\FrameSep\fbox{##1}
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed}
\makeatother


% *** CODE LISTING ***
\usepackage{listings}
\usepackage{color} 
\usepackage{textcomp}
\usepackage{float}
\floatstyle{plaintop}
\newfloat{Code}{H}{myc}
\definecolor{lbcolor}{rgb}{0.95,0.95,0.95} 

\usepackage{graphicx} 

\lstset{
	language=sh,
	keywordstyle=\bfseries\ttfamily\color[rgb]{0,0,1},
	identifierstyle=\ttfamily,
	commentstyle=\color[rgb]{0,0,1},
	stringstyle=\ttfamily\color[rgb]{0,0,1},
	showstringspaces=false,
	basicstyle=\scriptsize\ttfamily,
	numberstyle=\tiny,
	numbers=left,
	stepnumber=1,
	numbersep=5pt,
	tabsize=2,
	breaklines=true,
	prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
	frame=tb,
	breakatwhitespace=false,
	aboveskip={\baselineskip},
	belowskip={\baselineskip},
	columns=fixed,
	upquote=true,
	extendedchars=true,
	xleftmargin=2em,
	backgroundcolor=\color[rgb]{1,1,1},
	captionpos=b,  
}

\definecolor{add_green}{rgb}{0,0.6,0}

\lstdefinelanguage{diff}{
  %morecomment=[f][\color{blue}]{@@},     % group identifier
  morecomment=[f][\color{red}]-,         % deleted lines 
  morecomment=[f][\color{add_green}]+,       % added lines (dark green)
  morecomment=[f][\color{magenta}]{---}, % Diff header lines (must appear after +,-)
  morecomment=[f][\color{magenta}]{+++},
}

\newcommand*\rot{\rotatebox{90}}

\lstnewenvironment{code}[1][]%
{
   \noindent
   \minipage{\linewidth} 
   \vspace{0.5\baselineskip}
   \lstset{basicstyle=\ttfamily\footnotesize,frame=single,#1}}
{\endminipage}


% *** MATH PACKAGES ***
%
\usepackage{amsmath}


\usepackage[utf8]{inputenc}
\usepackage{subfigure}
\usepackage{framed}

\usepackage{xspace}
\newcommand*{\eg}{e.g.,\@\xspace}
\newcommand*{\ie}{i.e.,\@\xspace}
\newcommand{\secref}[1]{Section~\ref{#1}\xspace}
\newcommand{\figref}[1]{Figure~\ref{#1}\xspace}
\newcommand{\tabref}[1]{Table~\ref{#1}\xspace}
\newcommand{\listingref}[1]{Listing~\ref{#1}\xspace}

\usepackage{relsize}
\renewcommand*{\UrlFont}{\ttfamily\smaller\relax}
\usepackage{newunicodechar}
%for awkward characters: 
\DeclareUnicodeCharacter{00A0}{~}

\hyphenation{chan-ges}

\begin{document}

\title{FEVER: An Approach to Analyze Feature-Oriented Changes and Artefact Co-Evolution in Highly Configurable Systems}
\titlerunning{FEVER: Feature Evolution and Artefact Co-evolution}

\author{Nicolas Dintzner \and Arie van Deursen \and Martin Pinzger}

\institute{Nicolas Dintzner \and Arie van Deursen
\at Software Engineering Research Group,
Delft University of Technology,
Delft, Netherlands,\\
\email{N.J.R.Dintzner@tudelft.nl},
\email{Arie.vanDeursen@tudelft.nl}
\and
Martin Pinzger
\at Software Engineering Research Group,
University of Klagenfurt,
Klagenfurt, Austria,\\
\email{martin.pinzger@aau.at}
}

\maketitle

\begin{abstract}
The evolution of highly configurable systems is known to be a challenging task. Thorough understanding of configuration options
their relationships, and their implementation in various types of artefacts (variability model, mapping, and implementation) is required
to avoid compilation errors, invalid products, or dead code.

	Recent studies focusing on co-evolution of artefacts detailed feature-oriented change scenarios, describing how
related artefacts might change over time. However, relying on manual analysis of commits, such work
do not provide the means to obtain quantitative information on the frequency of described scenarios nor
information on the exhaustiveness of the presented scenarios for the evolution of a large scale system.

	In this work, we propose FEVER and its instantiation for the Linux kernel. 
FEVER extracts detailed information on changes in variability models (KConfig files), assets (preprocessor based C code), and mappings (Makefiles).
We apply this methodology to the Linux kernel and build a dataset comprised of 15 releases of the kernel history.

	We performed an evaluation of the FEVER approach by manually inspecting the data and compared it
with commits in the system's history. The evaluation shows that FEVER accurately captures feature related changes
for more than 85\% of the 810 manually inspected commits.
We use the collected data to reflect on occurrences of co-evolution in practice. 
Our analysis shows that complex co-evolution scenarios occur in every studied release but are not among
the most frequent change scenarios, as they only occur for 8 to 13\% of the evolving features.
Moreover, only a minority of developers working on a given release 
will make changes to all artefacts related to a feature (between 10\% and 13\% of authors).

	While our conclusions are derived from observations on the evolution of the Linux kernel, we believe that they may have
implications for tool developers as well as guide further research in the field of co-evolution of artefacts.
\end{abstract}

\keywords{highly variable systems, co-evolution, feature, variability}
\begin{acknowledgements}
This publication was supported by the Dutch national program COMMIT and carried out as part of the Allegio project 
under the responsibility of the Embedded Systems Innovation group of TNO.
\end{acknowledgements}



\input{introduction}
\input{background}
\input{approach}
\input{experiment}
\input{scenarios}
\input{co-evolution}
\input{related}
\input{conclusion}


%\section*{Acknowledgements}


\bibliographystyle{apalike}
\bibliography{refs}

%

\end{document}
